---
title: Quality report for `r qcData['File Name']`
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(reshape2)
knitr::opts_chunk$set(dpi=300)
```

# Basic statistics

`r knitr::kable(summaryData, align="lr")`


# Per base sequence quality boxplot

```{r, echo=FALSE,out.width="100%", fig.width=10,fig.height=5, fig.cap=""}
# Quality per cycle box plot 
boxplotData<-rqcData$CycleQualityHist 
boxplotData$cycle<-as.numeric(boxplotData$cycle)
maxQuality<-max(DEFAULT_MAXQUALITY,max(boxplotData$ymax))
plotData_even <- subset(boxplotData, cycle %% 2 == 0)
plotData_odd <- subset(boxplotData, cycle %% 2 == 1)
p<-ggplot(boxplotData) +
  # Draw background of low quality region
  geom_rect(data=plotData_even, aes(xmin = cycle - 0.5, xmax = cycle + 0.5, ymin = 0,ymax = 20),
    fill = rgb(1, 0, 0, alpha = 0.2)
  )+
  geom_rect(data=plotData_odd, aes(xmin = cycle - 0.5, xmax = cycle + 0.5, ymin = 0,ymax = 20),
    fill = rgb(1, 0, 0, alpha = 0.5)
  )+
  # Draw background of mid quality region
  geom_rect(data=plotData_even, aes(xmin = cycle - 0.5, xmax = cycle + 0.5, ymin = 20,ymax = 28),
    fill = rgb(1, 1, 0, alpha = 0.2)
  )+
  geom_rect(data=plotData_odd, aes(xmin = cycle - 0.5, xmax = cycle + 0.5, ymin = 20,ymax = 28),
    fill = rgb(1, 1, 0, alpha = 0.5)
  )+
  # Draw background of high quality region
  geom_rect(data=plotData_even, aes(xmin = cycle - 0.5, xmax = cycle + 0.5, ymin = 28,ymax = maxQuality),
    fill = rgb(0, 1, 0, alpha = 0.2)
  )+
  geom_rect(data=plotData_odd, aes(xmin = cycle - 0.5, xmax = cycle + 0.5, ymin = 28,ymax = maxQuality),
    fill = rgb(0, 1, 0, alpha = 0.5) 
  )+
  # Draw the line across median (REMOVED)
  # geom_line(aes(x = cycle, y = middle), color = "blue", linewidth = 1) +
  # Draw the lower whisker (min to Q1)
  geom_segment(aes(x = cycle, xend = cycle, y = ymin, yend = lower), 
               color = "black") +
  # Draw the lower whisker end at min
  geom_segment(aes(x = cycle - 0.3, xend = cycle + 0.3, y = ymin, yend = ymin), 
               color = "black") +

  # Draw the upper whisker (Q3 to max)
  geom_segment(aes(x = cycle, xend = cycle, y = upper, yend = ymax), 
               color = "black") +
  # Draw the upper whisker end at max
  geom_segment(aes(x = cycle - 0.3, xend = cycle + 0.3, y = ymax, yend = ymax), 
               color = "black") +
  # Draw the box (Q1 to Q3)
  geom_rect(aes(xmin = cycle - 0.3, xmax = cycle + 0.3, ymin = lower, ymax = upper), 
            fill = "yellow", color = "black") +
  # Draw the median line
  geom_segment(aes(x = cycle - 0.3, xend = cycle + 0.3, y = middle, yend = middle), 
               color = "red", linewidth = 1) +
  # Draw the line across average
  geom_line(data=rqcData$ReadQuality,aes(x = as.numeric(cycle), y = quality), color = "blue", linewidth = 0.5) +  
  # Add labels and theme
  labs(x = "Cycle", y = "Quality", title = "Per cycle Quality") +
  theme_classic()+
#  theme(axis.text.x = element_text(angle = -45))+
  scale_y_continuous(name="Quality", breaks=seq(0,maxQuality,2), labels=seq(0,maxQuality,2), limits=c(0, maxQuality),expand = c(0, 0))+
  scale_x_continuous(name="Position in read (bp)", breaks=c(1,seq(5,max(boxplotData$cycle),5)), labels=c(1,seq(5,max(boxplotData$cycle),5)), limits=c(0.5,max(boxplotData$cycle)+0.5),expand = c(0, 0))
#  scale_x_continuous(name="Position in read (bp)", breaks=1:max(boxplotData$cycle), labels=1:max(boxplotData$cycle), limits=c(0.5,max(boxplotData$cycle)+0.5),expand = c(0, 0))
  
print(p)
```

  

# Per base sequence content


```{r, echo=FALSE,out.width="100%", fig.width=10,fig.height=5, fig.cap=""}
# Nucleotide content per cycle line plot  (before filtering)
ntContent <- qcData$NucleotideContent %>% select(-c("GC"))
ntContent$POSITION=seq_along(ntContent$A)
meltDf<-reshape2::melt(ntContent,id.vars="POSITION", variable.name = "NT", value.name = "PERCENT")
if(showFiltered){  

	# Nucleotide content per cycle line plot  (after filtering)
	filteredNtContent <- qcData$FilteredNucleotideContent %>% select(-c("GC"))
	filteredNtContent$POSITION=seq_along(filteredNtContent$A)
	filteredMeltDf<-reshape2::melt(filteredNtContent,id.vars="POSITION", variable.name = "NT", value.name = "PERCENT")

	meltDf$Filtered="NO"
	filteredMeltDf$Filtered="YES"
	combinedDf<-rbind(meltDf,filteredMeltDf)
} else {
	combinedDf<-meltDf
}


maxY<-max(0.5,max(combinedDf$PERCENT))
rectData<-subset(ntContent, POSITION %% 2 == 1)
p<-ggplot(data=combinedDf) +
  geom_rect(data=rectData, aes(xmin = POSITION - 0.5, xmax = POSITION + 0.5, ymin = 0,ymax = maxY), fill = rgb(0.7, 0.7, 0.7, alpha = 0.5))+
  labs(x = "Cycle") +
  scale_color_manual(breaks = c("A","C","G","T","N"), labels=c("%A","%C","%G","%T","%N"), values=c(A="red",C="blue",G="green",T="black",N="purple"))+
  theme_classic()+
  theme(axis.line.x=element_line(linewidth = 0.5, colour = "lightgrey"),axis.line.y.right=element_line(linewidth = 0.5, colour = "lightgrey"),axis.ticks.y.right = element_line(linewidth = 0.5, colour = "lightgrey"),panel.grid.major.y = element_line(linewidth = 0.5, colour = "lightgrey"))+
  scale_y_continuous(name="Nucleotide content", limits=c(0, maxY),sec.axis = dup_axis(name = NULL, labels = NULL),expand = c(0, 0.001))+
  scale_x_continuous(name="Position in read (bp)", breaks=c(1,seq(5,max(ntContent$POSITION),5)), labels=c(1,seq(5,max(ntContent$POSITION),5)), limits=c(0.5,max(ntContent$POSITION)+0.5),expand = c(0, 0))

if(showFiltered){  
  p<-p+geom_line(aes(x=POSITION,y=PERCENT, color=NT, linetype=Filtered))
} else {
  p<-p+geom_line(aes(x=POSITION,y=PERCENT, color=NT))
}   

print(p)
```


# Sequence Duplication Levels

```{r, echo=FALSE,out.width="100%", fig.width=10,fig.height=5, fig.cap=""}
# Duplication rate
duplicationBreaks<-c(1:9,10,50,100,500,1000,5000,10000,Inf)
dupIntervals<-data.frame(INTERVAL=cut(duplicationBreaks[-length(duplicationBreaks)], breaks = duplicationBreaks, right = FALSE))
duplicationNames<-c(as.character(1:9),">10",">50",">100",">500",">1k",">5k",">10k")
duplicationNames<-factor(duplicationNames,levels = unique(duplicationNames))
plotData<-rqcData$DuplicationFrequencies
dupRateData<-plotData %>%
  mutate(INTERVAL = cut(occurrence, breaks = duplicationBreaks, right = FALSE)) %>%
  group_by(INTERVAL) %>%
  summarise(TOTAL = sum(percentage),DEDUPLICATED = sum(deduplicated))

dupRateData<-dupIntervals %>% dplyr::left_join(dupRateData,by=c("INTERVAL")) %>% mutate(TOTAL = ifelse(is.na(TOTAL), 0, TOTAL)) %>% mutate(DEDUPLICATED = ifelse(is.na(DEDUPLICATED), 0, DEDUPLICATED))

meltDf<-reshape2::melt(dupRateData, id.vars="INTERVAL", variable.name = "SEQUENCES", value.name = "PERCENTAGE")
oddColumns <-dupRateData[c(TRUE,FALSE), ]

p<-ggplot(data=meltDf,aes(x=INTERVAL,y=PERCENTAGE,color=SEQUENCES, group=SEQUENCES)) +
  geom_line(linewidth=1)+  
  geom_tile(data=oddColumns, aes(x = INTERVAL, y = 50, width=1, height=100),fill = rgb(0.7, 0.7, 0.7, alpha = 0.5), inherit.aes = FALSE)+  
  geom_line(linewidth=1)+    
  labs(title = paste0("Percent of sequences remaining if deduplicated: ",rqcData$DuplicationRate ,"%")) +
  scale_x_discrete(name="Number of repetitions", breaks=levels(meltDf$INTERVAL), labels=duplicationNames,expand = c(0, 0.5))+
  scale_y_continuous(name="Percentage", breaks=seq(0,100,10), labels=seq(0,100,10),sec.axis = dup_axis(name = NULL, labels = NULL),expand = c(0, 0))  +
  scale_color_manual(name="", breaks=levels(meltDf$SEQUENCES), labels=levels(meltDf$SEQUENCES), values=c(TOTAL="blue",DEDUPLICATED="red"))  +
  guides(colour = guide_legend(position = "inside"))+
  theme_classic()+
  theme(axis.line.x=element_blank(),axis.line.y.right=element_line(linewidth = 0.5, colour = "lightgrey"),axis.ticks.y.right = element_line(linewidth = 0.5, colour = "lightgrey"),panel.grid.major.y = element_line(linewidth = 0.5, colour = "lightgrey"),legend.position.inside = c(0.85, 0.9),legend.background = element_blank(), legend.key = element_blank())

  
print(p)
```

# Overrepresented sequences

`r if(overrepresented1) knitr::kable(overrepresentedDataSummary, align="lrr")  else "No overrepresented sequences found"`


# Adapter content

```{r, eval=adapter1, echo=FALSE,out.width="100%", fig.width=10,fig.height=5, fig.cap=""}
if(adapter1){
	# Cumulative percentage of occurrences of the adapter at each position: 
	plotData<-data.frame(POSITION=seq_along(adapter_cumsum),CUMSUM=adapter_cumsum)
	rectData<-subset(plotData, POSITION %% 2 == 1)
	maxY<-100
	p<-ggplot(data=plotData) +
	  geom_rect(data=rectData, aes(xmin = POSITION - 0.5, xmax = POSITION + 0.5, ymin = 0,ymax = maxY), fill = rgb(0.7, 0.7, 0.7, alpha = 0.5))+
	  geom_line(aes(x=POSITION,y=CUMSUM))+
	  labs(title = paste0("Percentage of reads with adapter: ",round(seqAdapterPercent*100,2),"%")) +
	  theme_classic()+
	  theme(axis.line.x=element_line(linewidth = 0.5, colour = "lightgrey"),axis.line.y.right=element_line(linewidth = 0.5, colour = "lightgrey"),axis.ticks.y.right = element_line(linewidth = 0.5, colour = "lightgrey"),panel.grid.major.y = element_line(linewidth = 0.5, colour = "lightgrey"))+
	  scale_y_continuous(name="Cumulative percent (adapter found)", breaks=seq(0,maxY,10), limits=c(0, maxY),sec.axis = dup_axis(name = NULL, labels = NULL),expand = c(0, 0.1))+
	  scale_x_continuous(name="Position in read (bp)", breaks=c(1,seq(5,max(plotData$POSITION),5)), labels=c(1,seq(5,max(plotData$POSITION),5)), limits=c(0.5,max(plotData$POSITION)+0.5),expand = c(0, 0))


	print(p)
}
```


`r if(adapter1) knitr::kable(AdapterStats, align="lr") else "Adapter not found"`


```{r, filtered-heading, eval=showFiltered,echo=FALSE}
	if(showFiltered){
	knitr::asis_output("# Estimated filtering statistics")
	}
```

`r if (showFiltered) knitr::kable(filteredData, align="lr")`


```{r, average-quality-heading, eval=showFiltered,echo=FALSE}
	if(showFiltered){
		knitr::asis_output("# Average per base sequence quality")
	}
```

```{r,eval=showFiltered,echo=FALSE,out.width="100%", fig.width=10,fig.height=5, fig.cap=""}
if(showFiltered){
	# Per cycle Quality (before and after filtering)
	meltDf<-reshape2::melt(qcData$Quality,id.vars="POSITION", variable.name = "GROUP", value.name = "QUALITY")
	maxQuality<-max(DEFAULT_MAXQUALITY,max(meltDf$QUALITY))
	plotData_even <- subset(meltDf, POSITION %% 2 == 0)
	plotData_odd <- subset(meltDf, POSITION %% 2 == 1)
	p<-ggplot(data=meltDf, aes(x=POSITION,y=QUALITY, linetype=GROUP)) +
	  # Draw background of low quality region
	  geom_rect(data=plotData_even, aes(xmin = POSITION - 0.5, xmax = POSITION + 0.5, ymin = 0,ymax = 20),
	    fill = rgb(1, 0, 0, alpha = 0.2),color = NA
	  )+
	  geom_rect(data=plotData_odd, aes(xmin = POSITION - 0.5, xmax = POSITION + 0.5, ymin = 0,ymax = 20),
	    fill = rgb(1, 0, 0, alpha = 0.5),color = NA
	  )+
	  # Draw background of mid quality region
	  geom_rect(data=plotData_even, aes(xmin = POSITION - 0.5, xmax = POSITION + 0.5, ymin = 20,ymax = 28),
	    fill = rgb(1, 1, 0, alpha = 0.2),color = NA
	  )+
	  geom_rect(data=plotData_odd, aes(xmin = POSITION - 0.5, xmax = POSITION + 0.5, ymin = 20,ymax = 28),
	    fill = rgb(1, 1, 0, alpha = 0.5),color = NA
	  )+
	  # Draw background of high quality region
	  geom_rect(data=plotData_even, aes(xmin = POSITION - 0.5, xmax = POSITION + 0.5, ymin = 28,ymax = maxQuality),
	    fill = rgb(0, 1, 0, alpha = 0.2),color = NA
	  )+
	  geom_rect(data=plotData_odd, aes(xmin = POSITION - 0.5, xmax = POSITION + 0.5, ymin = 28,ymax = maxQuality),
	    fill = rgb(0, 1, 0, alpha = 0.5),color = NA
	  )+  
	  geom_line()+
	  #labs(x = "Position", y = "Average Quality", title = "Per cycle Quality") +
	  labs(title = "Per cycle Quality") +
	  #ylim(0, maxQuality)+  
	  theme_classic()+
	  guides(linetype=guide_legend(title="Filtered",override.aes=list(fill=NA)))+
	  theme(legend.background=element_blank())+
	  scale_y_continuous(name="Quality", breaks=seq(0,maxQuality,2), labels=seq(0,maxQuality,2), limits=c(0, maxQuality),expand = c(0, 0))+
	#  scale_x_continuous(name="Position in read (bp)", breaks=1:max(meltDf$POSITION), labels=1:max(meltDf$POSITION), limits=c(0.5,max(meltDf$POSITION)+0.5),expand = c(0, 0))
	  scale_x_continuous(name="Position in read (bp)", breaks=c(1,seq(5,max(meltDf$POSITION),5)), labels=c(1,seq(5,max(meltDf$POSITION),5)), limits=c(0.5,max(meltDf$POSITION)+0.5),expand = c(0, 0)) 

	print(p)
}
```

